### 2020-07-14 reproduction/Tabs/tab2.tex
added reproduction results; comments in reproduction.rmd

% latex table generated in R 3.5.1 by xtable 1.8-4 package
% Mon Jul 13 19:13:01 2020
\begin{tabular}{cccr@{--}lr@{--}l}
  \hline
  & \multicolumn{2}{c}{Estimate (\%)} & \multicolumn{4}{c}{95\% credible interval}\\ Rule & original & reproduction & \multicolumn{2}{c}{original} & \multicolumn{2}{c}{reproduction}\\ \hline
100\% & 34 & 34 & 27 & 39 & 27 & 39 \\ 
  90\% & 36 & 36 & 29 & 42 & 29 & 42 \\ 
  5'80\%-3'100bp & 35 & 35 & 28 & 41 & 28 & 41 \\ 
  80\% & 40 & 40 & 33 & 46 & 33 & 46 \\ 
  70\% & 42 & 42 & 35 & 49 & 35 & 49 \\ 
  60\% & 42 & 42 & 33 & 50 & 33 & \textbf{\color{red} 49} \\ 
   \hline
\end{tabular}